---
title: "Peer-graded Assignment: Course Project 1"  
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
#prevent output in scientific notation
options(scipen=999)
```
Code for reading in the dataset and/or processing the data  
```{r, echo=TRUE}
library(lattice)


setwd('C://MOOC//ReproducibleResearch//')
a <- read.csv('activity.csv')
#calculate number of steps per day
r <-aggregate(a$steps, by=list(a$date), FUN=sum, na.rm=TRUE)
colnames(r) <- c("date", "StepsPerDay")
#calculate mean value for each interval
i <-aggregate(a$steps, by=list(a$interval), FUN=mean, na.rm=TRUE)
colnames(i) <- c("interval", "meansteps")
```

Histogram of the total number of steps taken each day 
```{r, echo=TRUE}
hist(r$StepsPerDay,
     main="Total number of steps taken each day",
     xlab="steps",
     breaks=nrow(r))
```


```{r, echo=TRUE}
me <- mean(r$StepsPerDay[r$StepsPerDay!=0])
med <- median(r$StepsPerDay[r$StepsPerDay!=0])
```
Mean `r me` and median number `r med` of steps taken each day  

Time series plot of the average number of steps taken 
```{r, echo=TRUE}
plot(i$interval, i$meansteps, main='Average number of steps taken in an interval', xlab='interval', ylab='Average number of steps', type='l')

```
 
```{r, echo=TRUE}
IntAvgMax <- subset(i, i$meansteps==max(i$meansteps))
```
The 5-minute interval that, on average, contains the maximum number of steps is `r IntAvgMax$interval`

```{r, echo=TRUE}
#missing values
nmv <- length(a$steps[a$steps=='NA'])
#impute mean value for interval if NA
imp <- a
imp <- merge(a, i, by="interval")
imp$steps <- ifelse(is.na(imp$steps), imp$meansteps, imp$steps)
#calculate steps per day
imp.r <-aggregate(imp$steps, by=list(imp$date), FUN=sum, na.rm=TRUE)
colnames(imp.r) <- c("date", "StepsPerDay")
imp.me <- mean(imp.r$StepsPerDay[imp.r$StepsPerDay!=0])
imp.med <- median(imp.r$StepsPerDay[imp.r$StepsPerDay!=0])
```
The total number of missing steps, steps=NA, in the dataset is `r nmv` 
Missing step values have been replaced by the average number of steps for that interval.  
After imputation, Mean `r imp.me` and median number `r imp.med` of steps taken each day  
Histogram of the total number of steps taken each day after missing values are imputed  
```{r, echo=TRUE}
hist(imp.r$StepsPerDay,
     main="Total number of steps taken each day (after imputation)",
     xlab="steps",
     breaks=nrow(imp.r))
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends 
```{r, echo=TRUE}
a$dayname <- weekdays(as.Date(a$date))
a$weekend <- "FALSE"
a$weekend <- ifelse(a$dayname %in% c("Saturday", "Sunday"), "weekend", "weekday")
wk <- subset(a, a$weekend=="weekday")
i.wk <-aggregate(wk$steps, by=list(wk$interval, wk$weekend), FUN=mean, na.rm=TRUE)
colnames(i.wk) <- c("interval", "weekend", "meansteps")
wkend <-subset(a, a$weekend=="weekend")
i.wkend <-aggregate(wkend$steps, by=list(wkend$interval, wkend$weekend), FUN=mean, na.rm=TRUE)
colnames(i.wkend) <- c("interval", "weekend", "meansteps")

p <- rbind(i.wk, i.wkend)
p$weekend <- as.factor(p$weekend)
xyplot(p$meansteps ~ p$interval | p$weekend,
       data = p,
       type = "l",
       xlab = "interval",
       ylab = "mean number of steps",
       main = "Average number of steps in an interval",
       layout=c(1,2))
```
